---
title: "Fish counts data curation"
author: "Allison Horst"
date: "8/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(janitor)
```


Curating the 'fish counts' LTER data into a perfect, tidy, ggplot-able CSV file. 

**Info:** https://portal.edirepository.org/nis/mapbrowse?scope=knb-lter-sbc&identifier=17&revision=newest

**Citation:** Reed D. 2018. SBC LTER: Reef: Kelp Forest Community Dynamics: Fish abundance. Environmental Data Initiative. https://doi.org/10.6073/pasta/dbd1d5f0b225d903371ce89b09ee7379. Dataset accessed 8/30/2019.

NOTE: "Count values were given a NA (-99999) if: (1) data were not collected for a particular species, (2) a datasheet was lost or (3) poor underwater visibility prevented fish from being surveyed" (from: https://portal.edirepository.org/nis/metadataviewer?packageid=knb-lter-sbc.17.33)

1. Get the raw data (updating all non-observed -99999's to NA):
```{r}

fish_counts_raw <- read_csv("fish_counts_raw.csv", na = "-99999")

```

2. Some basic cleaning (already in great shape): 
```{r}

fish_counts <- fish_counts_raw %>% 
  janitor::clean_names() %>% 
  purrr::map(tolower) %>% 
  as.data.frame()

```

3. Write CSV
```{r}

write_csv(fish_counts, "fish_counts.csv") # NOT THE FINAL ONE

```

4. Some further curation testing (trying out different fish subsets, etc.)
```{r}
# Subset fish
# Find the most commonly observed fish (by year and species) - will be more interesting than a bunch of 0's

fish_pivot <- fish_counts %>% 
  drop_na(count) %>% 
  mutate(count = as.numeric(count)) %>% 
  group_by(common_name) %>% 
  summarize(
    total = sum(count, na.rm = TRUE)
    ) %>% 
  arrange(-total)

top_ten <- fish_pivot$common_name[1:10]

# Try: senorita, blacksmith, black surfperch, rock wrasse, garibaldi

fish_subset <- fish_counts %>% 
  filter(common_name %in% c("senorita", "blacksmith", "black surfperch", "rock wrasse", "garibaldi")) %>% 
  filter(year %in% 2016:2018) %>% 
  filter(month == 7) %>% 
  drop_na(count) %>% 
  mutate(count = as.numeric(count)) %>% 
  group_by(year, site, common_name) %>% 
  summarize(
    total_count = sum(count)
  ) %>% 
  ungroup()

# This is the final curated version (fish_counts_curated.csv)
write_csv(fish_subset, "fish_counts_curated.csv")

# Make a graph of the numbers by location, year & species?
# Probably want to work on this a bit more ()

ggplot(fish_subset, aes(x = year, y = total_count, fill = common_name)) +
  geom_col(position = "dodge") +
  facet_wrap(~site)

```

