---
title: "Kelp counts data curation"
author: "Allison Horst"
date: "9/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(writexl) # install.packages("writexl")
library(here)
```


Curating the 'annual_kelp_all_years.csv' LTER data into an Excel worksheet with multiple tabs (stored as kelp_counts.xlsx). 

**Info:** https://portal.edirepository.org/nis/mapbrowse?scope=knb-lter-sbc&identifier=18&revision=newest

**Citation:** Reed D. 2018. SBC LTER: Reef: Kelp Forest Community Dynamics: Abundance and size of Giant Kelp (Macrocystis Pyrifera), ongoing since 2000. Environmental Data Initiative. https://doi.org/10.6073/pasta/dea56e02127161194626f97c8b6118e8. Dataset accessed 9/05/2019.

1. Get the raw data, clean and filter
```{r}

data_raw <- read_csv("annual_kelp_all_years.csv")

## clean!
data_clean <- data_raw %>% 
  janitor::clean_names() %>% 
  mutate(site = str_to_lower(site), sp_code = str_to_lower(sp_code))


# filter as we do
kelp_subset <- data_clean %>% 
  filter(year %in% 2016:2018, 
         month == 7) %>%
  drop_na(fronds) %>% 
  mutate(fronds = as.numeric(fronds)) %>% 
  mutate(year = as.numeric(year)) %>% # Adding to make year numeric (was character)
  group_by(year, month, site, common_name) %>% 
  summarize(
    total_fronds = sum(fronds)
  ) %>%
  ungroup() %>% 
  select(-common_name)

unique(kelp_subset$site) 

```

2. Split it up by site (will put data for each of 11 sites in a separate Excel worksheet, within the same workbook)
```{r}

kelp_split <- kelp_subset %>% 
  group_by(site) %>% 
  dplyr::group_split()

```

3. Save to separate sheets!
```{r}

# Just get the names & arrange (in order of list appearance)
site_names <- data_clean %>%
  select(site) %>%
  arrange(site)

# Set the names of the kelp_split list items
names(kelp_split) <- tolower(unique(site_names$site))

# Write to an Excel file: 
write_xlsx(kelp_split, here("curation", "kelp_counts_curated.xlsx"))
write_xlsx(kelp_split, here("data", "kelp_fronds.xlsx"))
```

