---
title: "lobster counts data curation"
author: "Allison Horst"
date: "8/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(janitor)
library(writexl)
```


Curating the 'lobster counts' data into a perfect, tidy, ggplot-able CSV file. 

**Info:**

**Citation:**
NOTE: "Count values were given a NA (-99999) if: (1) data were not collected for a particular species, (2) a datasheet was lost or (3) poor underwater visibility prevented lobster from being surveyed" (from: https://portal.edirepository.org/nis/metadataviewer?packageid=knb-lter-sbc.17.33)

1. Get the raw data (updating all non-observed -99999's to NA):
```{r}

lobster_counts_raw <- read_csv("curation/lobster_counts.csv", na = "-99999")

```

2. Some basic cleaning (already in great shape), and create subsets to save: 
```{r}

lobster_counts <- lobster_counts_raw %>% 
  janitor::clean_names() %>% 
  mutate(site = str_to_lower(site))

## will become lobsters2 to read in second
lobster_uncount_full <- lobster_counts %>%
  uncount(count) %>%
  select(-area, -num_ao)


## will become lobsters to read in first
lobster_uncount_subset <- lobster_uncount_full %>%
  filter(year <= 2016) 

```

3. rbind rows to demo skipping them during workshop
```{r}
lines_to_add <- tribble(
  ~year, ~month, ~date, ~site, ~transect, ~replicate, ~size_mm, 
  "Description: Lobster size, abundance and fishing pressure (SB coast)", 
    "", "", "", "", "", "", 
  "Link: https://portal.edirepository.org/nis/mapbrowse?scope=knb-lter-sbc&identifier=77&revision=newest", 
  "", "", "", "", "", "",   
  "Citation: Reed D. 2019. SBC LTER: Reef: Abundance, size and fishing effort for California Spiny Lobster (Panulirus interruptus), ongoing since 2012. Environmental Data Initiative. [doi](https://doi.org/10.6073/pasta/a593a675d644fdefb736750b291579a0).", 
  "", "", "", "", "", "", 
  "", "", "", "", "", "", "",
)

## rbind to objects
lobster_uncount_full <- rbind(lines_to_add, lobster_uncount_full)
lobster_uncount_subset <- rbind(lines_to_add, lobster_uncount_subset)

```


5. Save 2 files for pivot_tables chapter
```{r}

# These are the final curated version (lobster_counts_curated.csv)
write_xlsx(lobster_uncount_subset, "data/lobsters.xlsx")
write_xlsx(lobster_uncount_full, "data/lobsters2.xlsx")

# Make a graph of the numbers by location, year & species?
# Probably want to work on this a bit more ()

ggplot(lobster_subset, aes(x = year)) +
  geom_bar() +
  facet_wrap(~site)

```

