---
title: "Substrate benthic cover data curation"
author: "Allison Horst"
date: "9/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(tidyverse)
library(janitor)
library(writexl)
```


Data: https://portal.edirepository.org/nis/mapbrowse?scope=knb-lter-sbc&identifier=15&revision=newest

Citation: Reed D. 2018. SBC LTER: Reef: Kelp Forest Community Dynamics: Cover of sessile organisms, Uniform Point Contact. Environmental Data Initiative. https://doi.org/10.6073/pasta/f4ea94a612ef3dc0e14579dfb6298467. Dataset accessed 9/28/2019.

**Relevant variables:**

SUBSTRATE_TYPE:
- B = Bedrock
- BL = Large boulders (> 1 m diameter)
- BM = Medium boulders (50 - 100 cm)
- BS = Small boulders (25 - 50 cm)
- C = Cobble rock (< 25 cm)
- S = Sand, greater than 2.5 cm deep
- SH = Shell debris, broken
- SS = Sand shallow, less than 2.5 cm deep

PERCENT_COVER: 
"Percent cover of the species noted as determined by a uniform point contact method (80 point per 40x2 meter transect) where multiple layers of species at a given point are accounted for."

**Curation goals:** Make an Excel file containing a single Excel worksheet that contains non-data header rows (so we can use skip = n) and some empty cells (or -9999 indicators) that we can instead read in as NA.

Reading in data & some cleaning: 
```{r}
substrate <- read_csv("substrate_benthic_cover.csv") %>% 
  clean_names() %>% 
  purrr::map(tolower) %>% 
  as.data.frame()
```

Then I'll make some rows to add to the top of it. Note: the rep("",8) in each line ensures that they each have 9 columns, so that it can be bound to the 'substrate' df, which also has 9 columsn. 
```{r}

# Create the 4 lines:
line_one <- c("Substrate cover dataset, SBC LTER", rep("",8))
line_two <- c("Source: https://portal.edirepository.org/nis/mapbrowse?scope=knb-lter-sbc&identifier=15&revision=newest", rep("",8))
line_three <- c("Accessed: 9/28/2019", rep("",8))
line_four <- rep("",9)
line_five <- names(substrate)
line_six <- c(-9999, -9999, -9999, "carp", 1, 20, "i","b",-9999)
line_seven <- c(2000, 9, -9999, "carp", 1, 20, "o","b", -9999)

# Combine them all:
extra_lines <- rbind(line_one, line_two, line_three, line_four, line_five, line_six, line_seven)

# Make sure the variable names are the same in each: 
colnames(extra_lines) <- names(substrate)

```

Then combine them: 
```{r}

substrate_cover <- rbind(extra_lines, substrate) # Cool. 
# Notice that I'll need to exclude column names later. 

```

And create a CSV (excluding column names): 
```{r}
write_xlsx(substrate_cover, "substrate_cover_curated.xlsx", col_names = FALSE)
```

Done! Now this is a single Excel worksheet that has header rows and several -9999 entries right up top to demonstrate reading Excel worksheets. 